\textbf{How open are hybrid journals included in transformative agreements?}

### 2024-03-18 analysis/manuscript.tex
arxiv version

% number lines
% \usepackage{lineno}
% \linenumbers

@@

\abstract{The ongoing controversy surrounding transformative agreements, which aim to transition journal publishing to full open access, highlight the need for large-scale studies assessing the uptake of open access in hybrid journals. This includes evaluating the extent to which transformative agreements enabled open access. By combining publicly available data from various sources, including cOAlition S Journal Checker, Crossref, and OpenAlex, this study presents a novel approach that analyses over 700 agreements and nine million journal articles published in more than 11.000 hybrid journals. Estimates suggest a strong growth in open access between 2018 and 2022 from 4.3\% to 15\%. In 2022, 58\% of hybrid open access was enabled by transformative agreements. This trend was largely driven by the three commercial publishers Elsevier, Springer Nature, and Wiley, but the open access uptake varied substantially across journals, publishers, disciplines, and country affiliations. In particular, comparing the developments in the OECD and BRICS areas revealed different publication trends relative to hybrid open access. In conclusion, estimates suggest that current levels of implementation of transformative agreements is insufficient to bring about a large-scale transition to full open access.}

@@

% QSS wants numbered sections
%\setcounter{secnumdepth}{2}

@@

For over two decades, hybrid open access journal publishing, which makes
some articles openly available while others remain behind a paywall, has
been discussed as a means for transitioning the subscription system to
full open access (Prosser, 2003). The idea was that when journals
increasingly publish open access articles, they could reduce revenues
from subscriptions, while libraries and funders could change their
funding models and shift expenditures from subscription to open access.
However, initial approaches, mainly based on publication fees, also
called article processing charges (APCs), did not contribute
substantially to a large open access uptake. In 2009, Springer reported
open access to 1\% of articles in hybrid journals (Dallmeier-Tiessen et
al., 2010). Other studies have also recorded low uptake. In 2011, only
1-2\% of articles were open access (Björk, 2012), increasing to around
4\% between 2011 and 2013 (Laakso \& Björk, 2016).

@@

By the end of 2023, many transformative agreements had been implemented,
but the interim outcomes were mixed. The ESAC Transformative Agreement
Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}},
the largest source of disclosure, recorded more than 800 transformative
agreements, resulting in up to 900.000 open access articles published in
both fully open access and hybrid journals according to the accompanying
ESAC Market Watch\footnote{\url{https://esac-initiative.org/market-watch/}}.
Library consortia reported increased open access volume, streamlined
payment and monitoring procedures, as well as extensive utilisation of
open access options by the researchers they serve (Marques \& Stone,
2020; Parmhed \& Säll, 2023; Pinhasi et al., 2020). The ongoing
standardisation of transformative agreements has contributed to improved
transparency in terms of contracts and publisher-provided article
metadata (Marques et al., 2019; Pinhasi et al., 2021). However, with the
growing trend toward transformative agreements, continued reliance on
big deals is perceived as problematic, because it perpetuates market
concentration (Butler et al., 2023; Shu \& Larivière, 2023). Whether
transformative agreements lead to reduced pricing remains uncertain
(Borrego, 2023) and a substantial transition of hybrid journals towards
full open access could not be observed (Matthias et al., 2019; Momeni et
al., 2021). The focus on large commercial publishers might also increase
inequality (Ross-Hellauer et al., 2022), because transformative
agreements' focus on pay to publish mainly targets institutions from
high-income countries, furthering a questionable journal prestige
culture (Babini et al., 2022). Besides, an editorial-board resignation
raised concerns that transformative agreements might encourage
publishers to maximize journal publication volume ``without regard to
quality'' (Rasmussen, 2023).

The controversies surrounding hybrid open access and transformative
agreements have led to varying policy conclusions. For instance, the
Association of Swedish Higher Education Institutions (Sveriges
universitets- och högskoleförbund, SUHF) recommended only supporting
agreements for publishing in fully open access journals \footnote{\url{https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787}}.
Likewise, most cOAlition S funders will discontinue financial support
for transformative agreements by the end of 2024 (Liverpool, 2023). The
consortium also removed the majority of hybrid journals from its
Transformative Journal program in 2023 due to publishers' failure to
meet self-defined open access growth targets (Brainard, 2023). By
contrast, the German DEAL consortium announced a five-year
transformative agreement with Elsevier starting in 2024 and also renewed
its contracts with Springer Nature and Wiley until the end of 2028.
Similarly, the Colombia Consortium signed the first transformative
agreements in Latin America (Muñoz-Vélez et al., 2024).

Despite these controversies around transformative agreements as a means
of transitioning journal publishing to full open access, there is
limited evidence available on the uptake of open access in hybrid
journals, and the extent to which this can be attributed to
transformative agreements. Previous studies have focused on specific
countries (Haucap et al., 2021; Huang et al., 2020; Pölönen et al.,
2020; Taubert et al., 2023; Wenaas, 2022) or publisher portfolios
(Bakker et al., 2024; Fraser et al., 2023; Jahn et al., 2022; Momeni et
al., 2023; Pieper \& Broschinski, 2018; Schmal, 2024), while large-scale
studies relied on self-reported agreement data (Moskovkin et al., 2022),
or used APC pricing lists (Shu \& Larivière, 2023). In particular, data
availability is a limiting factor when studying the impact of
transformative agreements (Bakker et al., 2024), because bibliometric
databases, even though many allow the retrieval of open access articles
in hybrid journals, do not directly attribute them to specific
transformative agreements. Likewise, article-level open access invoicing
data are only partly available (Jahn et al., 2022).

@@

A limitation of using the Journal Checker Tool and its underlying
publicly available data dump to study the development of transformative
agreements over time is that expired transformative agreements are
constantly removed. To address this, four different snapshots were
safeguarded and combined for this study: self-archived versions from
July 2021, July 2022, and May 2023, as well as the most current dump
downloaded on 11 December 2023. This ensured that transformative
agreements, which ended from 2021 onwards, were included, representing
the majority of transformative agreements. Overall, the four combined
Transformative Agreement Data dumps used in this study contained 729 of
869 agreements listed in the ESAC registry by December 2023.

The Transformative Agreement Data dumps link agreements to journals
represented by journal names and the ISSN. After mapping ISSN variants
to the corresponding linking ISSN (ISSN-L), as provided by the ISSN
International Centre, journals were associated with publishers using the
ESAC ID, a unique identifier for transformative agreements in the ESAC
Transformative Agreement Registry. Furthermore, journal subjects
according to the All Science Journal Classification (ASJC) were added
from the Scopus journal source list as of August 2023.

@@

After identifying hybrid journals included in transformative agreements,
article metadata were retrieved from the Crossref November 2023 database
snapshot for the five-year period from 2018 to 2022, according to the
issued date, representing the earliest known publication date. Because
Crossref metadata lacked information to distinguish between original
research articles, including reviews, and other types of journal
content, which are often not covered by transformative agreements
(Borrego et al., 2021), only articles published in regular issues
indicated by non-numeric pagination were included. Furthermore, paratext
recognition was applied to exclude non-scholarly journal content such as
table of contents.

@@

After retrieving the article metadata, the publication volume, including
open access, was calculated for each journal. To improve the
identification of hybrid journals, journals with an open access
proportion above 95\% were excluded. This step allowed removing
additional 241 fully open access journals.

Affiliation metadata about corresponding authors are crucial for the
planning and evaluation of transformative agreements because they are
considered responsible for arranging open access publication (Borrego et
al., 2021; Geschuhn \& Stone, 2017; Schimmer et al., 2015). For this
study, country and institutional affiliations were retrieved from
OpenAlex. However, because corresponding author affiliation was only
available for 54\% of the investigated articles in OpenAlex, this study
focused on first authors and their affiliations instead; approximately
90\% of the investigated articles had first author affiliation metadata
in OpenAlex, which is a much larger proportion than previoulsy reported
for the October 2022 snapshot (Zhang et al., 2024). First authors
typically contribute the most to a paper and are often considered being
lead authors (Larivière et al., 2016). Related studies also assumed
first authors as a proxy to examine open access payments and
transformative agreements (Haucap et al., 2021; Shu \& Larivière, 2023;
Zhang et al., 2022).

@@

Throughout this mostly automated data gathering and analysis process,
Tidyverse tools (Wickham et al., 2019) for the R programming language (R
Core Team, 2020) were used. The resulting data are openly available
through an R data package, hoaddata (Jahn, 2023). Following Marwick et
al. (2018), hoaddata contains not only the datasets used in the data
analysis. It also includes code used to compile the data by connecting
it to the cloud-based Google Big Query data warehouse, where scholarly
big data from Crossref, OpenAlex and Unpaywall were imported, using
bigrquery (Wickham \& Bryan, 2023). To increase computational
reproducibility, data aggregation through hoaddata was automatically
carried out using GitHub Actions, a continuous integration service.

@@

Figure \ref{fig:results_overview}B highlights that the majority of
hybrid open access was made available through transformative agreements
in 2021 and 2022, contributing 58\% of the total open access article
volume in 2022. However, there was also a notable increase in open
access provision through other means, presumably publication fees being
not invoiced through transformative agreements, from 4.1\% (n = 62,625)
in 2018 to 6.3\% (n = 105,896) in 2022.

@@

However, there are notable differences between the three large
publishers. Although Elsevier published the largest volume of articles
(n = 2,770,826, 34\%), it recorded a comparably low number of open
access articles, including those that are associated with transformative
agreements. In contrast, Springer Nature and Wiley provided open access
to a larger proportion of their articles (13\% of Springer Nature
articles and 15\% of Wiley articles were open access), leading to higher
open access market share (23\% Springer Nature resp. 23\% Wiley). This
difference between Elsevier on the one hand and Springer Nature and
Wiley on the other can be attributed to transformative agreements, as
the latter made the majority of their open access articles available
through such deals (Springer Nature 57\% resp. Wiley 55\%).

@@

Between 2018 and 2022, high-income countries almost exclusively
dominated hybrid open access publishing through transformative
agreements. During this period, first authors affiliated with
institutions from Organisation for Economic Co-operation and Development
(OECD) member countries published 602,050 open access articles in hybrid
journals, representing 81\% of the investigated open access articles.
This disparity between OECD nations and other countries becomes even
more evident when considering open access through transformative
agreements, as 310,712 of 328,957, or 94\% of open access articles were
associated with such agreements.

@@

Moreover, this study presents varying levels of open access uptake
through transformative agreements across journals, which can be
attributed to the alignment of authors' affiliations and the
availability of such agreements at their institutions. In line with
previous research findings (Butler et al., 2023; Jahn et al., 2022;
Wenaas, 2022), high uptake rates were observed across hybrid journals in
Social Sciences and Humanities. However, it is important to emphasise
that these hybrid journals do not encompass the entire field. For
example, Khanna et al. (2022) found that approximately 60\% of journals
utilizing the open-source publishing platform Open Journal Systems (OJS)
fall within the Social Sciences and Humanities. In these fields, the
majority of full open access journals are so-called ``Diamond OA
journals'', which do not charge publication fees (Bosman, Frantsvåg, et
al., 2021). Rather, this result can be more accurately attributed to the
substantial proportion of authors from high-income countries who publish
in these hybrid journals included in transformative agreements,
particularly when journals are of local relevance, such as by belonging
to national societies or regional research scope.

@@

This large-scale study provides the first empirical evidence of the
influence of transformative agreements on the transition of hybrid
journals to full open access. However, several limitations needs to be
acknowledged. From a data perspective, estimations of open access
through transformative agreements were established by linking first
author affiliations with publicly available agreement data from
cOAlition S, and not through invoicing data, which is usually not
shared. Moreover, this study is unable to account for the various types
of transformative agreements, particularly with regard to article types
and caps that limit the number of open access articles covered.
Furthermore, assessing the quality of the OpenAlex and Crossref data
used, particularly in terms of affiliations and article types, by
combining them with established bibliometric databases such as Scopus
and Web of Science was beyond the scope of this analysis (Visser et al.,
2021). Overall, the methodology is designed in such a way that the
results underestimate rather than overestimate the adoption of open
access through transformative agreements. In addition, the data used in
this study are openly available, along with the code used for this
analysis.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Brainard_2023}{}}%
Brainard, J. (2023). {`Transformative'} journals get booted for
switching to open access too slowly. \emph{Science}.
\url{https://doi.org/10.1126/science.adj3282}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-Goodin_2023}{}}%
Rasmussen, K. B. (2023). Interview with {Robert {`Bob'} E. Goodin}.
\emph{Tidskrift För Politisk Filosofi}.
\url{https://www.politiskfilosofi.se/fulltext/2023-2/pdf/TPF_2023-2_interview_with_robert_bob_e_goodin.pdf}

>>>>

% number lines
 \usepackage{lineno}
 \linenumbers

@@

\abstract{The ongoing controversy surrounding transformative agreements, which aim to transition journal publishing to full open access, highlight the need for large-scale studies assessing the impact of these agreements on hybrid open access. By combining publicly available data from various sources, including cOAlition S Journal Checker, Crossref, and OpenAlex, this study presents a novel approach that analyses over 700 agreements. Results suggest a strong growth in open access between 2018 and 2022 from 4.3\% to 15\%. During this period, 11,189 hybrid journals provided open access to 742,369 out of 8,146,958 articles, representing a five-year open access proportion of 9.1\%. Authors who could make use of transformative agreements at the time of publication contributed 328,957 open access articles. In 2022, 143,615 out of 249,511 open access articles in hybrid journals (58\%) were enabled by transformative agreements. This trend was largely driven by the three commercial publishers Elsevier, Springer Nature, and Wiley, but the open access uptake varied substantially across journals, publishers, disciplines, and country affiliations. In particular, the OECD and BRICS areas revealed different publication trends. In conclusion, this study suggests that current levels of implementation of transformative agreements is insufficient to bring about a large-scale transition to full open access.}

@@

% QSS wants numbered sections
\setcounter{secnumdepth}{2}

@@

For over two decades, hybrid open access journal publishing, which makes
some articles openly available while others remain behind a paywall, has
been discussed as a means for transitioning the subscription system to
full open access (Prosser, 2003). The idea was that when journals
increasingly publish open access articles, they could reduce revenues
from subscriptions, while libraries and funders could change their
funding models and shift expenditures from subscription to open access.
However, initial approaches, mainly based on publication fees, also
called article processing charges (APCs), did not contribute
substantially to a large open access uptake. In 2009, the publisher
Springer reported that 1\% of its articles in hybrid journals were open
access (Dallmeier-Tiessen et al., 2010). Other studies have also
recorded low uptake. In 2011, only 1-2\% of articles were open access
(Björk, 2012), increasing to around 4\% between 2011 and 2013 (Laakso \&
Björk, 2016).

@@

By the end of 2023, many transformative agreements had been implemented,
but the interim outcomes were mixed. The ESAC Transformative Agreement
Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}},
the largest resource for library consortia to publicly share their
agreements, recorded more than 800 transformative agreements. These
agreements resulted in the publication of up to 900.000 open access
articles published in both fully open access and hybrid journals,
according to the accompanying ESAC Market Watch\footnote{\url{https://esac-initiative.org/market-watch/}}.
Library consortia reported increased open access volume, streamlined
payment and monitoring procedures, as well as extensive utilisation of
open access options by the researchers they serve (Marques \& Stone,
2020; Parmhed \& Säll, 2023; Pinhasi et al., 2020). The ongoing
standardisation of transformative agreements has contributed to improved
transparency in terms of contracts and publisher-provided article
metadata (Marques et al., 2019; Pinhasi et al., 2021). However, with the
growing trend toward transformative agreements, continued reliance on
big deals is perceived as problematic, because it perpetuates market
concentration (Butler et al., 2023; Shu \& Larivière, 2023). Whether
transformative agreements lead to reduced pricing remains uncertain
(Borrego, 2023) and a substantial transition of hybrid journals towards
full open access could not be observed (Matthias et al., 2019; Momeni et
al., 2021). The focus on large commercial publishers might also increase
inequality (Ross-Hellauer et al., 2022), because transformative
agreements' focus on pay to publish mainly targets institutions from
high-income countries, furthering a questionable journal prestige
culture (Babini et al., 2022). Besides, an editorial-board resignation
raised concerns that transformative agreements might encourage
publishers to maximize journal publication volume ``without regard to
quality'' (Rasmussen, 2023).

The controversies surrounding hybrid open access and transformative
agreements have led to varying policy conclusions. For instance, the
Association of Swedish Higher Education Institutions (Sveriges
universitets- och högskoleförbund, SUHF) recommended only supporting
agreements for publishing in fully open access journals by 2026
\footnote{\url{https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787}}.
Likewise, most cOAlition S funders will discontinue financial support
for transformative agreements by the end of 2024 (Liverpool, 2023).
cOAlition S also removed the majority of hybrid journals from its
Transformative Journal program in 2023 due to publishers' failure to
meet self-defined open access growth targets (Brainard, 2023). By
contrast, the German DEAL consortium announced a five-year
transformative agreement with Elsevier starting in 2024 and also renewed
its contracts with Springer Nature and Wiley until the end of 2028.
Similarly, the Colombia Consortium signed the first transformative
agreements in Latin America (Muñoz-Vélez et al., 2024).

Despite these controversies around transformative agreements as a means
of transitioning journal publishing to full open access, there is
limited evidence available on the uptake of open access in hybrid
journals, and the extent to which this can be attributed to
transformative agreements. Previous studies have focused on specific
countries (Brayman et al., 2024; Haucap et al., 2021; Huang et al.,
2020; Pölönen et al., 2020; Taubert et al., 2023; Wenaas, 2022) or
publisher portfolios (Bakker et al., 2024; Fraser et al., 2023; Jahn et
al., 2022; Momeni et al., 2023; Pieper \& Broschinski, 2018; Schmal,
2024), while large-scale studies relied on self-reported agreement data
(Moskovkin et al., 2022), or used APC pricing lists (Shu \& Larivière,
2023). In particular, data availability is a limiting factor when
studying the impact of transformative agreements (Bakker et al., 2024),
because bibliometric databases, even though many allow the retrieval of
open access articles in hybrid journals, do not directly attribute them
to specific transformative agreements. Likewise, article-level open
access invoicing and cost data are only partly available (Jahn et al.,
2022; Kramer, 2024).

@@

A limitation of using the Journal Checker Tool and its underlying
publicly available data dump to study the development of transformative
agreements over time is that expired transformative agreements are
constantly removed. To address this, four different snapshots were
safeguarded and combined for this study: self-archived versions from
July 2021, July 2022, and May 2023, as well as the most current dump
downloaded on 11 December 2023. This ensured that transformative
agreements, which ended from 2021 onwards, were included, representing
the majority of transformative agreements. Overall, the four combined
Transformative Agreement Data dumps used in this study contained 729 of
869 agreements listed in the ESAC Transformative Agreement Registry by
December 2023.

The Transformative Agreement Data dumps link agreements to journals
represented by journal names and the ISSN. After mapping ISSN variants
to the corresponding linking ISSN (ISSN-L), as provided by the ISSN
International Centre, journals were associated with publishers according
to the ESAC Transformative Agreement Registry. This reflects that some
portfolios may include imprints. Furthermore, journal subjects according
to the All Science Journal Classification (ASJC) were added from the
Scopus journal source list as of August 2023.

@@

After identifying hybrid journals included in transformative agreements,
article metadata were retrieved from the Crossref November 2023 database
snapshot for the five-year period from 2018 to 2022, according to the
issued date, representing the earliest known publication date. Because
Crossref metadata lacked information to distinguish between original
research articles, including reviews, and other types of journal
content, which are often not covered by transformative agreements
(Borrego et al., 2021), only articles published in regular issues
indicated by numeric pagination were included. Furthermore, paratext
recognition was applied to exclude non-scholarly journal content such as
table of contents.

@@

After retrieving the article metadata, the publication volume, including
open access, was calculated for each journal. To improve the
identification of hybrid journals, journals with an open access
proportion above 95\% were excluded. This step allowed removing
additional 134 fully open access journals. Together, these journals
published 8,565 articles between 2018 and 2022.

Affiliation metadata about corresponding authors are crucial for the
planning and evaluation of transformative agreements because they are
considered responsible for arranging open access publication (Borrego et
al., 2021; Geschuhn \& Stone, 2017; Schimmer et al., 2015). For this
study, country and institutional affiliations were retrieved from
OpenAlex. However, because corresponding author affiliation was only
available for 54\% of the investigated articles in OpenAlex, this study
focused on first authors and their affiliations instead; approximately
90\% of the investigated articles had first author affiliation metadata
in OpenAlex, which is a much larger proportion than previously reported
for the October 2022 snapshot (Zhang et al., 2024). First authors
typically contribute the most to a paper and are often considered being
lead authors (Larivière et al., 2016). Related studies also assumed
first authors as a proxy to examine open access payments and
transformative agreements (Haucap et al., 2021; Shu \& Larivière, 2023;
Zhang et al., 2022).

@@

Throughout this mostly automated data gathering and analysis process,
Tidyverse tools (Wickham et al., 2019) for the R programming language (R
Core Team, 2020) were used. The resulting data are openly available
through an R data package, hoaddata (Jahn, 2023). Following Marwick et
al. (2018), hoaddata contains not only the datasets used in the data
analysis. It also includes code used to compile the data by connecting
it to a cloud-based Google BigQuery data warehouse, where scholarly big
data from Crossref, OpenAlex and Unpaywall were made available, using
bigrquery (Wickham \& Bryan, 2023). To increase computational
reproducibility, data aggregation through hoaddata was automatically
carried out using GitHub Actions, a continuous integration service.

@@

Figure \ref{fig:results_overview}B highlights that the majority of
hybrid open access was made available through transformative agreements
in 2021 and 2022. In 2022, 143,615 out of 249,511 open access articles
were from eligible authors, representing 58\%. However, there was also a
notable increase in open access provision through other means,
presumably publication fees being not invoiced through transformative
agreements, from 4.1\% (n = 62,625) in 2018 to 6.3\% (n = 105,896) in
2022.

@@

However, there are notable differences between the three large
publishers. Although Elsevier published the largest volume of articles
(n = 2,770,826, 34\%), it recorded a comparably low number of open
access articles, including those that are associated with transformative
agreements. In contrast, Springer Nature and Wiley provided open access
to a larger proportion of their articles (13\% of Springer Nature
articles and 15\% of Wiley articles were open access), leading to higher
open access market share (23\% Springer Nature resp. 23\% Wiley). This
difference between Elsevier on the one hand and Springer Nature and
Wiley on the other can be attributed to transformative agreements, as
the latter made the majority of their open access articles available
through such deals (57\% Springer Nature resp. 55\% Wiley).

@@

Between 2018 and 2022, high-income countries almost exclusively
dominated hybrid open access publishing through transformative
agreements. To discern socio-economic differences, these countries were
grouped according to their membership in the Organisation for Economic
Co-operation and Development (OECD). During 2018 and 2022, first authors
affiliated with institutions from OECD member countries published
602,050 open access articles in hybrid journals, representing 81\% of
the investigated open access articles. This disparity between OECD
nations and other countries becomes even more evident when considering
open access through transformative agreements, as 310,712 of 328,957, or
94\% of open access articles were associated with such agreements.

@@

Moreover, this study presents varying levels of open access uptake
through transformative agreements across journals, which can be
attributed to the alignment of authors' affiliations and the
availability of such agreements at their institutions. In line with
previous research findings (Butler et al., 2023; Jahn et al., 2022;
Wenaas, 2022), high uptake rates were observed across Social Sciences
and Humanities hybrid journals. However, it is important to emphasise
that these hybrid journals do not encompass the entire field. For
example, Khanna et al. (2022) found that approximately 60\% of journals
utilising the open-source publishing platform Open Journal Systems (OJS)
fall within the Social Sciences and Humanities. In these fields, the
majority of full open access journals are so-called ``Diamond OA
journals'', which do not charge publication fees (Bosman, Frantsvåg, et
al., 2021). Rather, this result can be more accurately attributed to the
substantial proportion of authors from high-income countries who publish
in these hybrid journals included in transformative agreements,
particularly when journals are of local relevance, such as by belonging
to national societies or regional research scope.

@@

This large-scale study provides the first empirical evidence of the
influence of transformative agreements on the transition of hybrid
journals to full open access. However, several limitations need to be
acknowledged. From a data perspective, estimations of open access
through transformative agreements were established by linking first
author affiliations with publicly available agreement data from
cOAlition S, and not through invoicing data, which is usually not
shared. Moreover, this study is unable to account for the various types
of transformative agreements, particularly with regard to article types
and caps that limit the number of open access articles covered.
Furthermore, assessing the quality of the OpenAlex and Crossref data
used, particularly in terms of affiliations and article types, by
combining them with established bibliometric databases such as Scopus
and Web of Science was beyond the scope of this analysis (Visser et al.,
2021). Overall, the methodology is designed to underestimate, rather
than overestimate, the adoption of open access through transformative
agreements. In addition, the data used in this study are openly
available, along with the code used for this analysis.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Brainard_2023}{}}%
Brainard, J. (2023). {``Transformative''} journals get booted for
switching to open access too slowly. \emph{Science}.
\url{https://doi.org/10.1126/science.adj3282}

\leavevmode\vadjust pre{\hypertarget{ref-Jisc_2024}{}}%
Brayman, K., Devenney, A., Dobson, H., Marques, M., \& Vernon, A.
(2024). \emph{A review of transitional agreements in the {UK}}. Zenodo.
\url{https://doi.org/10.5281/zenodo.10787392}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Kramer_2024}{}}%
Kramer, B. (2024). \emph{Study on scientific publishing in {Europe} --
{Development}, diversity, and transparency of costs}. Publications
Office of the European Union. \url{https://doi.org/doi/10.2777/89349}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Goodin_2023}{}}%
Rasmussen, K. B. (2023). Interview with {Robert {``Bob''} E. Goodin}.
\emph{Tidskrift För Politisk Filosofi}.
\url{https://www.politiskfilosofi.se/fulltext/2023-2/pdf/TPF_2023-2_interview_with_robert_bob_e_goodin.pdf}
